The very dependent recursive structure of iterated parametricity in indexed form

### 2025-12-29 paper/paper.tex
paper: more prose

For $\nu$ an arity, that is $1$ in the case of predicates, families and augmented semi-simplicial sets or $2$ in the case of relations, correspondences and semi-cubical sets, such sequence of was called a $\nu$-set. Equations specifying $\nu$-sets were given in \cite{hr25}. Because of the dependency of $E_{n+1}$ into all $E_p$ for $p \leq n$, the family could not be expressed as a function. Instead, it was expressed as a stream whose $(n+1)$-th element $E_{n+1}$ was dependent on the bundle of all previous elements of the stream. Such a bundle was called $\Xto{m}{n}$ while the residual stream of all $E_{n}$, $E_{n+1}$, ..., all of them precisely dependent on the bundle $E:\Xto{m}{n}$ of previous $E_0$, ..., $E_{n-1}$, was called $\Xfrom{m}{n}{D}$. The type of $E_n$, also dependent on a bundle $D:\Xto{m}{n}$, was denoted $\Xcomp{m}{n}[D=\D]$, as shown on tables~\ref{tab:coind} and~\ref{tab:core}. When $n$ is $0$, $\Xto{m}{0}$ is degenerated, that is reduced to the canonical inhabitant $\kstar$ of $\unittype$. Then, $\Xfrom{m}{0}{\kstar}$ denotes the full stream $E_0, E_1, \ldots$.

@@

Notably, the definition of $\restrl$ relies on an equality expressing the commutation of the composition of two $\restrf$. The proof of this commutation is worth being made explicit, which we do in table~\ref{tab:coh} using proof-term notations. The proof is by recursion on the structure of $\framep$, $\layer$, and $\painting$. This is what $\cohf$ does using auxiliary proofs $\cohl$ and $\cohp$. Note that for a fixed constant $n$, relying on the evaluation rules of type theory, the coherence conditions degenerate to a reflexivity proof, so that the construction builds an effective sequence of types not mentioning coherences anymore.

TODO: say that the table are relative to ETT and that many type coercions are implicit.

>>>>

For $\nu$ an arity, that is $1$ in the case of predicates, families and augmented semi-simplicial sets or $2$ in the case of relations, correspondences and semi-cubical sets, such sequence of was called a $\nu$-set. Equations specifying $\nu$-sets were given in \cite{hr25}. Because of the dependency of $E_{n+1}$ into all $E_p$ for $p \leq n$, the family could not be expressed as a function. Instead, it was expressed as a stream whose $(n+1)$-th element $E_{n+1}$ was dependent on the bundle of all previous elements of the stream. Such a bundle was called $\Xto{m}{n}$ while the residual stream of all $E_{n}$, $E_{n+1}$, ..., all of them precisely dependent on the bundle $E:\Xto{m}{n}$ of previous $E_0$, ..., $E_{n-1}$, was called $\Xfrom{m}{n}{D}$. The type of $E_n$, also dependent on a bundle $D:\Xto{m}{n}$, was denoted $\Xcomp{m}{n}[D=\D]$, as shown on tables~\ref{tab:coind} and~\ref{tab:core}, using an type theory as language of discourse. When $n$ is $0$, $\Xto{m}{0}$ is degenerated, that is reduced to the canonical inhabitant $\kstar$ of $\unittype$. Then, $\Xfrom{m}{0}{\kstar}$ denotes the full stream $E_0, E_1, \ldots$.

@@

Notably, the definition of $\restrl$ relies on an equality expressing the commutation of the composition of two $\restrf$. The proof of this commutation is worth being made explicit, which we do in table~\ref{tab:coh} using proof-term notations\footnote{We however omit type conversions, as if working in extensional type theory. E.g. the definition of $\restrl$ implicitly use $\cohf$ so that the type of $\restrp$ gets its expected type.}. The proof is by recursion on the structure of $\framep$, $\layer$, and $\painting$. This is what $\cohf$ does using auxiliary proofs $\cohl$ and $\cohp$. Note that for a fixed constant $n$, relying on the evaluation rules of type theory, the coherence conditions degenerate to a reflexivity proof, so that the construction builds an effective sequence of types not mentioning coherences anymore.
